Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-11 shanghai.tex
Restructuring the section on finite character sheaves to relect the plethora of lemmata that are required.

\section{The category of geometrized characters}\label{section:category}

@@

%\begin{remark} \label{rem:CS-action}
The rule $(\gcs{L},\mu,\phi) \mapsto (\gcs{L},\phi)$ defines a forgetful functor from geometrized characters of $G$ to ($\ell$-adic) Weil sheaves on $G$ \cite{Deligne:Weil2}*{Def.~1.1.10 (i)}.
%
As remarked after \cite{Deligne:Weil2}*{Def.~1.1.10}, a Weil sheaf $(\gcs{L},\phi)$ on $G$ may be interpreted as a constructible $\ell$-adic sheaf $\gcs{L}$ on $\bG$ together with an action of the Weil group $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ on $\bG$. Since we will need this point of view later, we review that interpretation here. Setting 
\[
\varphi(\Frob{\Fq}^n) \ceq  \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi)
\]
defines, for every $w\in \Weil{\Fq}$, an isomorphism
\[
\varphi(w) : w_G^* \gcs{L}\to \gcs{L},
\]
such that
\[
\forall u,v\in \Weil{\Fq},\qquad \varphi(uv) = \varphi(v) \circ v_G^* \varphi(u).
\]
Note that $\phi = \varphi(\Frob{\Fq})$. 
If we further define $\varphi_1(w) \ceq (w_G)_*(\varphi(w)^{-1})$ then
\[
\varphi_1(w) : (w_G)_* \gcs{L}\to \gcs{L},
\]
is an isomorphism and
\[
 \forall u,v\in \Weil{\Fq},\qquad \varphi_1(uv) = \varphi_1(u) \circ (u_G)_* \varphi_1(v),
\]
which means the pair $(\gcs{L},\varphi_1)$ is almost an action $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ in the sense of \cite{SGA4.5}*{Expos\'e XIII,~1.1}. The pair $(\gcs{L},\varphi_1)$ fails to fulfill the requirements of \cite{SGA4.5}*{Expos\'e XIII,~1.1} only because $\Weil{\Fq}$ is profinite. This fact goes to the heart of the notion of a Weil sheaf and is one of the essential aspects of our definition of $\GC(G)$. 
%\end{remark}

@@

It is clear that $\trFrob{\cs{L}\otimes \cs{L'}} = \trFrob{\cs{L}} \cdot \trFrob{\cs{L'}}$
(pointwise multiplication of functions) and that $\trFrob{\cs{L}} = \trFrob{\cs{L'}}$
if $\cs{L} \iso \cs{L'}$ in $\GC(G)$. In this way we obtain a group homomorphism
\[
\cdef{\trFrob{G}} : \GCiso{G} \longrightarrow \Hom_\text{grp}(G(\Fq),\EEx),
\qquad \text{defined by}\qquad \cs{L} \mapsto \trFrob{\cs{L}},
\]
from the Picard group of geometrized characters of $G$ to the character group of $G(\Fq)$.

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gcs{L},\mu,\phi) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \GC(G) \to \GC(H)$ such that
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \GCiso{G} \rar{f^*} \arrow[swap]{d}{\trFrob{G}} & \GCiso{H} \dar{t_H} \\
  \Hom_\text{grp}(G(\Fq),\EEx) \rar & \Hom_\text{grp}(H(\Fq),\EEx)
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

Consider the short exact sequence in the category of smooth group
schemes, locally of finite type over $\Fq$, defining the component
group scheme for $G$:
\begin{equation}\label{eq:pi-2}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is an Abelian group scheme, smooth (in fact, etale) and locally of finite type over $\Fq$, 
it follows from Proposition~\ref{prop:pullback} that \eqref{eq:pi-1} defines a sequence of functors
\[
\begin{tikzcd}
\GC(0) \rar & \GC(\pi_0(G)) \rar & \GC(G) \rar & \GC(G^0) \rar & \GC(0)
\end{tikzcd}
\]
which induces, upon passage to Picard groups, a sequence of group homomorphisms 
\begin{equation}\label{eq:pi-1}
\begin{tikzcd}
\GCiso{0} \rar & \GCiso{\pi_0(G)} \rar & \GCiso{G} \rar & \GCiso{G^0} \rar & \GCiso{0}.
\end{tikzcd}
\end{equation}
It is easy to see that $\GCiso{0}$ is trivial. In this paper we will find $\GCiso{G^0}$ and $\GCiso{\pi_0(G)}$ under the hypothesis that $\pi_0(G)$ is geometrically finitely generated; 
the snake lemma will then allow us to show
\[
\GCiso{G} \iso \Hom_\text{grp}(G(\Fq),\EE^\times),
\]
canonically.

@@

\begin{proposition}\label{prop:basechange}
 The rule $(\gcs{L}, \mu, \varphi) \mapsto (\gcs{L}, \mu, \varphi\vert_{\Weil{\Fq}})$ 
 defines a monoidal functor $\GC(G) \to \GC(G_{k'})$. 
 Moreover, the following diagram commutes, 
 where the top map is induced by $\GC(G) \to \GC(G_{k'})$ 
 and the bottom map is induced by the norm $G(k') \rightarrow G(k)$:
\[
\begin{tikzcd}[column sep=45]
\GCiso{G} \rar{\text{ext'n of scalars}} \dar{\trFrob{G}} & \GCiso{G_{k'}} \dar{\trFrob{G\times_{\Spec{k}} \Spec{k'}}} \\
\Hom(G(k), \EEx) \rar{\Nm} & \Hom(G(k'), \EEx).
\end{tikzcd}
\]
\end{proposition}

@@

To define the category of finite geometrized characters we must work a bit harder.
Consider the following category \cdef{$C(G)$}:
an object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ an etale cover and a morphism of commutative group $\Fq$-schemes (an isogeny) 
for which the action of $\Gal(\bFq/\Fq)$ on the etale group scheme $\ker f$ is trivial, 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\alpha,T)$ 
where $\alpha$ is a morphism of schemes such that $f' = f\circ \alpha$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker f'$ on $V$ by $\psi \circ h$.

We put a tensor category structure on $C(G)$ as follows.
 Let $(f,\psi)$ and $(f',\psi')$ be objects in $C(G)$. 
 The category of isogenies to $G$ admits products: 
 write $f\times_G f' : H\times_G H' \to G$ for the product of $f : H\to G$ and $f' : H'\to G$. 
 Since $\ker(f\times_G f') = (\ker f)\times (\ker f')$, 
 the tensor product of the representations $\psi : \ker f\to \Aut(V)$ and $\psi' : \ker f'\to \Aut(V')$ 
 is the representation $\psi \otimes\psi' :  \ker(f\times_G f') \to \Aut(V\otimes V')$. 
 This allows us to set $(f,\psi)\otimes(f',\psi') \ceq (f\times_G f' ,\psi\otimes\psi')$. 
 The same fact makes sense of $(f,\psi)\oplus(f',\psi')\ceq (f\times_G f' ,\psi\oplus\psi')$. 
 The dual of $(f,\psi)$ is $(f,\psi)^\vee \ceq (f,\psi^\vee)$ 
 where $\psi^\vee : \ker f \to \Aut(V^\vee)$ and $V^\vee = \Hom_\text{vec}(V,\EE)$. 
 The verification that these operations make $C(G)$ a tensor category is routine.

We say that $(\alpha,T)$ in $C(G)$ is a \cdef{weak isomorphism} if $T$ is an isomorphism of vector spaces. 
Let \cdef{$C(G)[W^{-1}]$} be the localization of $C(G)$ at weak isomorphisms.
Now, let \cdef{$C_1(G)$} be the subcategory of $C(G)$ such that $\psi$ is a one-dimensional representation
and let \cdef{$\GCf(G)$} be the localisation of $C_1(G)$ at weak isomorphisms. 
This is the rigid symmetric monoidal category of \cdef{finite geometrized characters} of $G$.

@@

 Now consider the localization $C(G)[W^{-1}]$ of $C(G)$ at weak isomorphisms.
 Morphisms in $C(G)[W^{-1}]$ are simple to describe because category $C(G)$ 
 admits a calculus of left fractions\footnote{See {\tt http://ncatlab.org/nlab/show/calculus+of+fractions}.} for weak isomorphisms: 
 every isomorphism in $C(G)$ is a weak isomorphism; 
 push-outs exist in $C(G)$ and the push-out of a weak isomorphism is a weak isomorphism; 
 and weak isomorphisms enjoy the left-cancellation property. 
 It follows that a morphism in $C(G)[W^{-1}]$ is the equivalence class of 
  \[
   \begin{tikzcd}
   (f,\psi) \arrow{r}{(\alpha_1,T_1)} & (f_1,\psi_1) & \arrow[swap]{l}{(\alpha'_1,T'_1)} (f',\psi')
   \end{tikzcd}
  \]
 where $(\alpha',T') : (f,\psi) \to (f',T')$ is any map in $C(G)$ 
 and $(\alpha'_1,T'_1) : (f_1,\psi_1) \to (f',\psi')$ is a weak isomorphism.
% this morphism is commonly denoted by $(\alpha_1,T_1)^{-1}\circ (\alpha'_1,T'_1)$. 
 %The equivalence relation identifies $(\alpha_1,T_1)^{-1}\circ (\alpha'_1,T'_1)$ 
 % with $(\alpha_2,T_2)^{-1}\circ (\alpha'_2,T'_2)$ if there are maps 
 %$(f_1,\psi_1)\to (f_0,\psi_0)$ and $(f_2,\psi_2)\to (f_0,\psi_0)$ making the following diagram commute
 %\[
 %  \begin{tikzcd}
 %  & (f_1,\psi_1) \arrow{d} & \\
 %  (f,\psi) \arrow{ur}{(\alpha_1,T_1)} \arrow{dr}{(\alpha_2,T_2)} & (f_0,\psi_0)  & \arrow[swap]{ld}{(\alpha'_2,T'_2)} \arrow[swap]{lu}{(\alpha'_1,T'_1)} (f',\psi')\\
 %  & (f_2,\psi_2) \arrow{u} & \\
 %  \end{tikzcd}
 %\]
 %where $(f',\psi') \to (f_0,\psi_0)$ is a weak isomorphism.

 Observe that $L$ takes weak isomorphisms in $C(G)$ to isomorphisms in $\Loc(G)$. 
 To see this, let $(h,T) : (f,\psi)\to (f',\psi')$ be a weak isomorphism, 
 where $f:H\to G$ and $f':H'\to G$ as above. 
 Since $T$ is an isomorphism, we may replace $V'$ with $V$ and $T$ with $\id_V$, without loss of generality. 
 Then the equivariance condition on $T$ becomes $\psi' = \psi\circ h$ on the finite etale group scheme $\ker f'$. 
 Since $f' = f \circ \alpha$ and $f$ and $f'$ are etale covers, 
 it follows that $\alpha$ is also an etale cover. 
 \Clifton{Is that right? If $f' = f \circ \alpha$ and $f$ and $f'$ are etale covers, 
 does it follow that $\alpha$ is also an etale cover? (Etale: Yes. Finite? Surjective?) If not, then I must modify the definition of weak isomorphism accordingly.} 
 It follows that $\alpha : H'\to H$ admits a section $\alpha' : H \to H'$, 
 which provides the inverse to $L(\alpha,T) : L(f,\psi) \to L(f',\psi')$, above. 
 \Clifton{Is it clear?}
 It is also clear that if $L(\alpha,T)$ is an isomorphism of local systems, 
 then $T$ must be an isomorphism of vector spaces, 
 whence $(\alpha,T)$ is a weak isomorphism: $W$ is saturated.

@@

  Every morphism of group schemes $G'\to G$ 
  induces a monoidal functor $C(G) \to C(G')$ by base-change along $G'\to G$ 
  ({\it cf.}, Proposition~\ref{prop:pullback}) as follows. 
  Let $m : G'\to G$ be a morphism of group schemes. 
  Let $(f,\psi)$ be an object in $C(G)$.

  indeed, the pull-back of $(f,\psi)$ along any morphism $m : G'\to G$ of group schemes 
  can be read off the short exact sequence obtained by pulling back along $m$.

  Taking the case when $m$ is the multiplication morphism $m : G\times G\to G$, 
  and using the tensor category structure above, 
  we see that the pull-back of $(f,\psi)$ along $m$ is canonically isomorphic, in $C_1(G\times G)$, to $(f,\psi)\boxtimes(f,\psi)$.
  \Clifton{I need to work on this diagram and explain better what's going on here. Under construction.}
  \[
   \begin{tikzcd}
V & \Aut(V) & \arrow[swap]{l}{\psi} \ker f \arrow{r}  & H \arrow{r}{f} & G \\
V & \Aut(V) \arrow[equal]{u} & \arrow[swap]{l}{\psi_{m_H}} \ker (f\times f) \arrow{r} \arrow{u}{m_H} & \arrow{d} H\times H \arrow{r}{f\times f} \arrow{u}{m_H} & G\times G \arrow{u}{m} \\
V\otimes V & \Aut(V\otimes V) & \arrow[swap]{l}{\psi\otimes \psi} \ker f \times \ker f \arrow{r}  & H\times H \arrow{r}{f\times f}  & G\times G \arrow[equal]{u} \\
   \end{tikzcd}
  \]

 If $(f,\psi)$ in an object in $C_1(G)$ then, as we have already seen, 
 $m^*L(f,\psi) \iso L(f,\psi) \boxtimes L(f,\psi)$ in $C_1(G\times G)$, canonically;
 let $\mu(f,\psi)$ be that isomorphism of $\ell$-adic sheaves on $G$,
 It follows immediately that the functor $L_G  \ceq  L_{W}\vert_{\GCf(G)} : \GCf(G) \to \GCb(G)$, 
 given on objects by $(f,\psi) \mapsto (L(f,\psi), \mu(f,\psi))$,
 is a full and faithfull monoidal functor, completing the proof.
\end{proof}

@@

We will see that the full subcategory $\GC_f(G) \subset \GC_0(G)$ of finite geometrized characters is a proper subcategory, in general. Indeed, it follows from our main result regarding geometrized characters, Theorem~\ref{thm:snake}, that the restriction of $\trFrob{G} : \GCiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx)$ to the subgroup $\GCfiso{G}$ is an isomorphism onto the subgroup of $\chi \in \Hom_\text{grp}(G(\Fq),\EEx)$ such that the image of $\chi$ is finite in $\EEx$.

@@

\begin{proposition}\label{prop:connected}
  If $G$ is a connected commutative algebraic group over $\Fq$ then 
  the full subcategories $\GCf(G) \hookrightarrow \GCb(G) \hookrightarrow \GC(G)$ 
  appearing in Section~\ref{sec:bounded} are equivalences 
  and
  \[
  \trFrob{G} : \GCiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx)
  \]
  is an isomorphism of groups.
\end{proposition}

@@

  To finish the proof of Proposition~\ref{prop:connected}, it suffices to observe that the group homomorphism
  \[
  \trFrob{G} : \GCiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx)
  \]
  is surjective because, as we have seen, 
  each character $\chi \in \Hom_\text{grp}(G(\Fq),\EEx)$ determines a finite geometrized character $(\Lang,\chi)$
  such that the trace of Frobenius of the geometrized character $B_G^* (L_G(\Lang,\chi))$ is $\chi$.
\end{proof}

>>>>

\section{The category of geometrized characters}\label{sec:category}

@@

%\begin{remark} \label{rem:CS-action}
The rule $(\gcs{L},\mu,\phi) \mapsto (\gcs{L},\phi)$ defines a forgetful functor from geometrized characters of $G$ to ($\ell$-adic) Weil sheaves on $G$ \cite{Deligne:Weil2}*{Def.~1.1.10 (i)}.

As mentioned after \cite{Deligne:Weil2}*{Def.~1.1.10}, a Weil sheaf $(\gcs{L},\phi)$ on $G$ may be interpreted as a constructible $\ell$-adic sheaf $\gcs{L}$ on $\bG$ together with an action of the Weil group $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ on $\bG$. Since we will need this point of view later, we review that interpretation here. Setting 
\[
\varphi(\Frob{\Fq}^n) \ceq  \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi)
\]
defines, for every $w\in \Weil{\Fq}$, an isomorphism
$\varphi(w) : w_G^* \gcs{L}\to \gcs{L}$, such that
$\varphi(uv) = \varphi(v) \circ v_G^* \varphi(u)$ for all $u,v\in \Weil{\Fq}$.
Note that $\phi = \varphi(\Frob{\Fq})$. 
If we further define $\varphi_1(w) \ceq (w_G)_*(\varphi(w)^{-1})$ then
$\varphi_1(w) : (w_G)_* \gcs{L}\to \gcs{L}$ is an isomorphism and
$\varphi_1(uv) = \varphi_1(u) \circ (u_G)_* \varphi_1(v)$ for all $u,v\in \Weil{\Fq}$.
This means the pair $(\gcs{L},\varphi_1)$ is almost an action $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ in the sense of \cite{SGA4.5}*{Expos\'e XIII,~1.1}. The pair $(\gcs{L},\varphi_1)$ fails to fulfill the requirements of \cite{SGA4.5}*{Expos\'e XIII,~1.1} only because $\Weil{\Fq}$ is profinite. This fact goes to the heart of the notion of a Weil sheaf and is one of the essential aspects of our definition of $\GC(G)$. 
%\end{remark}

@@

It is clear that $\trFrob{\cs{L}\otimes \cs{L'}} = \trFrob{\cs{L}} \cdot \trFrob{\cs{L'}}$
(pointwise multiplication of functions) and that $\trFrob{\cs{L}} = \trFrob{\cs{L'}}$
if $\cs{L} \iso \cs{L'}$ in $\GC(G)$. In this way we obtain a group homomorphism
\[
\cdef{\trFrob{G}} : \GCiso{G} \longrightarrow G(\Fq)^*,%\Hom_\text{grp}(G(\Fq),\EEx),
\qquad \text{defined by}\qquad \cs{L} \mapsto \trFrob{\cs{L}},
\]
from the Picard group of geometrized characters of $G$ to the character group of $G(\Fq)$.

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gcs{L},\mu,\phi) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \GC(G) \to \GC(H)$ such that
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \GCiso{G} \rar{f^*} \arrow[swap]{d}{\trFrob{G}} & \GCiso{H} \dar{t_H} \\
   G(\Fq)^* \rar & H(\Fq)^*
  %\Hom_\text{grp}(G(\Fq),\EEx) \rar & \Hom_\text{grp}(H(\Fq),\EEx)
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

Consider the short exact sequence in the category of smooth group
schemes, locally of finite type over $\Fq$, defining the component
group scheme for $G$:
\begin{equation}\label{eq:pi-2}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is an Abelian group scheme, smooth (in fact, etale) and locally of finite type over $\Fq$, 
it follows from Proposition~\ref{prop:pullback} that \eqref{eq:pi-1} defines a sequence of functors
\[
\begin{tikzcd}
\GC(0) \rar & \GC(\pi_0(G)) \rar & \GC(G) \rar & \GC(G^0) \rar & \GC(0)
\end{tikzcd}
\]
which induces, upon passage to Picard groups, a sequence of group homomorphisms 
\begin{equation}\label{eq:pi-1}
\begin{tikzcd}
\GCiso{0} \rar & \GCiso{\pi_0(G)} \rar & \GCiso{G} \rar & \GCiso{G^0} \rar & \GCiso{0}.
\end{tikzcd}
\end{equation}
It is easy to see that $\GCiso{0}$ is trivial. In this paper we will find $\GCiso{G^0}$ and $\GCiso{\pi_0(G)}$ under the hypothesis that $\pi_0(G)$ is geometrically finitely generated; 
the snake lemma will then allow us to show
\[
\GCiso{G} \iso G(\Fq)^*,%\Hom_\text{grp}(G(\Fq),\EE^\times),
\]
canonically.

@@

\begin{proposition}\label{prop:basechange}
 The rule $(\gcs{L}, \mu, \varphi) \mapsto (\gcs{L}, \mu, \varphi\vert_{\Weil{\Fq}})$ 
 defines a monoidal functor $\GC(G) \to \GC(G_{k'})$. 
 Moreover, the following diagram commutes, 
 where the top map is induced by $\GC(G) \to \GC(G_{k'})$ 
 and the bottom map is induced by the norm $G(k') \rightarrow G(k)$:
\[
\begin{tikzcd}[column sep=60]
\GCiso{G} \rar{\text{ext'n of scalars}} \dar{\trFrob{G}} & \GCiso{G_{k'}} \dar{\trFrob{G_{k'}}} \\
G(k)^*  \rar{\Nm} & G(k')^*
%\Hom(G(k), \EEx) \rar{\Nm} & \Hom(G(k'), \EEx).
\end{tikzcd}
\]
\end{proposition}

@@

To define the category of finite geometrized characters we must work a bit harder.
Consider the following category \cdef{$C(G)$}:
an object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ an isogeny (a finite, surjective etale morphism and a morphism of commutative group $\Fq$-schemes)
for which the action of $\Gal(\bFq/\Fq)$ on the etale group scheme $\ker f$ is trivial, 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\alpha,T)$ 
where $\alpha$ is a morphism of schemes such that $f' = f\circ \alpha$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker '$ on $V$ by $\psi \circ \alpha$.
\[
\begin{tikzcd}
V \arrow{d}{T} & \Aut(V) & \arrow[swap]{l}{\psi} \ker f \arrow[hook]{r} & H \arrow{r}{f} & \arrow[equal]{d} G\\
V' & \Aut(V) & \arrow[swap]{l}{\psi} \arrow[swap]{u}{\alpha\vert_{\ker f}} \ker f \arrow[hook]{r} & \arrow[swap]{u}{\alpha} H' \arrow{r}{f'} & G\\
\end{tikzcd}
\]
We say that $(\alpha,T)$ in $C(G)$ is a \cdef{weak isomorphism} if $\alpha$ is a covering and $T$ is an isomorphism of vector spaces.
\Clifton{I suspect $\alpha$ is a covering, for free. Is that right? 
If $f' = f \circ \alpha$ and $f$ and $f'$ are etale covers, 
does it follow that $\alpha$ is also an etale cover? (Etale: Yes. Finite? Surjective?) 
If so, then I can simplify the definition of weak isomorphism accordingly.}

We put a tensor category structure on $C(G)$ as follows.
Let $(f,\psi)$ and $(f',\psi')$ be objects in $C(G)$. 
The category of isogenies to $G$ admits products: 
write $f\times_G f' : H\times_G H' \to G$ for the product of $f : H\to G$ and $f' : H'\to G$. 
Since $\ker(f\times_G f') \iso (\ker f)\times (\ker f')$, canonically,
the tensor product of the representations $\psi : \ker f\to \Aut(V)$ and $\psi' : \ker f'\to \Aut(V')$ 
is the representation $\psi \otimes\psi' :  \ker(f\times_G f') \to \Aut(V\otimes V')$. 
This allows us to set $(f,\psi)\otimes(f',\psi') \ceq (f\times_G f' ,\psi\otimes\psi')$. 
The same fact makes sense of $(f,\psi)\oplus(f',\psi')\ceq (f\times_G f' ,\psi\oplus\psi')$. 
The dual of $(f,\psi)$ is $(f,\psi)^\vee \ceq (f,\psi^\vee)$ 
where $\psi^\vee : \ker f \to \Aut(V^\vee)$ and $V^\vee = \Hom_\text{vec}(V,\EE)$. 
The verification that these operations make $C(G)$ a tensor category is routine.

%
Let \cdef{$C(G)[W^{-1}]$} be the localization of $C(G)$ at weak isomorphisms.
Now, let \cdef{$C_1(G)$} be the subcategory of $C(G)$ such that $\psi$ is a one-dimensional representation
and let \cdef{$\GCf(G)$} be the localisation of $C_1(G)$ at weak isomorphisms. 
This is the rigid symmetric monoidal category of \cdef{finite geometrized characters} of $G$.

\begin{proposition}\label{prop:GCf-pull-back}
If $g : G'\to G$ is a morphism of group schemes then $(f,\psi) \mapsto (f_g,\psi_g)$ defines a monomial functor $g^* : \GCf(G) \to \GCf(G')$ where $f_g : H\times_G G'\to G'$ and $\psi_g : \ker f_g \to \Aut(V)$ are defined by the diagram below.
\[
\begin{tikzcd}
V & \Aut(V) & \arrow[swap]{l}{\psi} \ker f \arrow[hook]{r} & H \arrow{r}{f} &  G\\
  &  & \arrow[dashed]{ul}{\psi_g} \arrow[swap]{u}{(g_f)\vert_{\ker g_f}} \ker f_g \arrow[hook]{r} & \arrow[swap]{u}{g_f} H\times_G G' \arrow{r}{f_g} & \arrow{u}{g} G'.
\end{tikzcd}
\]
\end{proposition}

\begin{proposition}
Each $(f,\psi)\in \GCf(G)$ determines a canonical isomorphism $m^*(f,\psi) \to (f,\psi)\boxtimes(f,\psi)$ in $\GCf(G\times G)$.
 \end{proposition}

\begin{proof}
Before compare $m^*(f,\psi)$ and 
\end{proof}

Every morphism of group schemes $g: G'\to G$ 
defines a function $g^*: \obj C(G) \to \obj C(G')$ 
by $g^*(f,\psi) = (f_g,\psi_g)$ where
$f_g : H\times_G G'\to G'$ and $\psi_g : \ker f_g \to \Aut(V)$ are defined by the diagram below.
\[
\begin{tikzcd}
V & \Aut(V) & \arrow[swap]{l}{\psi} \ker f \arrow[hook]{r} & H \arrow{r}{f} &  G\\
  &  & \arrow[dashed]{ul}{\psi_g} \arrow[swap]{u}{(g_f)\vert_{\ker g_f}} \ker f_g \arrow[hook]{r} & \arrow[swap]{u}{g_f} H\times_G G' \arrow{r}{f_g} & \arrow{u}{g} G'.
\end{tikzcd}
\]
This definition is

@@

 Now consider the localization $C(G)[W^{-1}]$ of $C(G)$ at weak isomorphisms.
 Morphisms in $C(G)[W^{-1}]$ are simple to describe because category $C(G)$ 
 admits a calculus of left fractions\footnote{See {\tt http://ncatlab.org/nlab/show/calculus+of+fractions}.} for weak isomorphisms: 
 every isomorphism in $C(G)$ is a weak isomorphism; 
 push-outs exist in $C(G)$ and the push-out of a weak isomorphism is a weak isomorphism; 
 and weak isomorphisms enjoy the left-cancellation property. 
 It follows that a morphism in $C(G)[W^{-1}]$ is the equivalence class of 
  \[
   \begin{tikzcd}
   (f,\psi) \arrow{r}{(\alpha_1,T_1)} & (f_1,\psi_1) & \arrow[swap]{l}{(\alpha'_1,T'_1)} (f',\psi')
   \end{tikzcd}
  \]
 where $(\alpha',T') : (f,\psi) \to (f',T')$ is any map in $C(G)$ 
 and $(\alpha'_1,T'_1) : (f_1,\psi_1) \to (f',\psi')$ is a weak isomorphism.
 \Clifton{Reveal the commented-out details . . . I think I need them!}
% this morphism is commonly denoted by $(\alpha_1,T_1)^{-1}\circ (\alpha'_1,T'_1)$. 
 %The equivalence relation identifies $(\alpha_1,T_1)^{-1}\circ (\alpha'_1,T'_1)$ 
 % with $(\alpha_2,T_2)^{-1}\circ (\alpha'_2,T'_2)$ if there are maps 
 %$(f_1,\psi_1)\to (f_0,\psi_0)$ and $(f_2,\psi_2)\to (f_0,\psi_0)$ making the following diagram commute
 %\[
 %  \begin{tikzcd}
 %  & (f_1,\psi_1) \arrow{d} & \\
 %  (f,\psi) \arrow{ur}{(\alpha_1,T_1)} \arrow{dr}{(\alpha_2,T_2)} & (f_0,\psi_0)  & \arrow[swap]{ld}{(\alpha'_2,T'_2)} \arrow[swap]{lu}{(\alpha'_1,T'_1)} (f',\psi')\\
 %  & (f_2,\psi_2) \arrow{u} & \\
 %  \end{tikzcd}
 %\]
 %where $(f',\psi') \to (f_0,\psi_0)$ is a weak isomorphism.

 Observe that $L$ takes weak isomorphisms in $C(G)$ to isomorphisms in $\Loc(G)$. 
 To see this, let $(h,T) : (f,\psi)\to (f',\psi')$ be a weak isomorphism, 
 where $f:H\to G$ and $f':H'\to G$ as above. 
 Since $T$ is an isomorphism, we may replace $V'$ with $V$ and $T$ with $\id_V$, without loss of generality. 
 Then the equivariance condition on $T$ becomes $\psi' = \psi\circ h$ on the finite etale group scheme $\ker f'$. 
 Since $\alpha : H'\to H$ is an etale cover, it admits a section $\alpha' : H \to H'$, 
 which provides the inverse to $L(\alpha,T) : L(f,\psi) \to L(f',\psi')$, above. 
 It is also clear that if $L(\alpha,T)$ is an isomorphism of local systems, 
 then $T$ must be an isomorphism of vector spaces, 
 whence $(\alpha,T)$ is a weak isomorphism: $W$ is saturated.

@@

 If $(f,\psi)$ in an object in $C_1(G)$ then, as we have already seen, 
 $m^*L(f,\psi) \to L(f,\psi) \boxtimes L(f,\psi)$ by a weak isomorphism in $C_1(G\times G)$.
 This morphism in $C_1(G\times G)$ is not unique, but its image in $\GCf(G\times G)$ is . . .!

 let $\mu(f,\psi)$ be that isomorphism of $\ell$-adic sheaves on $G$,
 It follows immediately that the functor $L_G  \ceq  L_{W}\vert_{\GCf(G)} : \GCf(G) \to \GCb(G)$, 
 given on objects by $(f,\psi) \mapsto (L(f,\psi), \mu(f,\psi))$,
 is a full and faithfull monoidal functor, completing the proof.
\end{proof}

@@

We will see that the full subcategory $\GC_f(G) \subset \GC_0(G)$ of finite geometrized characters is a proper subcategory, in general. Indeed, it follows from our main result regarding geometrized characters, Theorem~\ref{thm:snake}, that the restriction of $\trFrob{G} : \GCiso{G} \to G(\Fq)^*$ to the subgroup $\GCfiso{G}$ is an isomorphism onto the subgroup of $\chi \in G(\Fq)^*$ such that the image of $\chi$ is finite in $\EEx$.

@@

\begin{proposition}\label{prop:connected}
  If $G$ is a connected commutative algebraic group over $\Fq$ then 
  the full subcategories $\GCf(G) \hookrightarrow \GCb(G) \hookrightarrow \GC(G)$ 
  appearing in Section~\ref{sec:bounded} are equivalences 
  and
  \[
  \trFrob{G} : \GCiso{G} \to G(\Fq)^*
  \]
  is an isomorphism of groups.
\end{proposition}

@@

  To finish the proof of Proposition~\ref{prop:connected}, it suffices to observe that the group homomorphism
  $
  \trFrob{G} : \GCiso{G} \to G(\Fq)^*
  $
  is surjective because each character $\chi \in G(\Fq)^*$ determines a finite geometrized character $(\Lang,\chi)$ such that the trace of Frobenius of the geometrized character $B_G^* (L_G(\Lang,\chi^{-1}))$ is $\chi$.
\end{proof}
